\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2012-10-25 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updated the manual

\section*{Authors}
The SPECFEM3D Cartesian package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\

Since then it has been developed and maintained by a development team: in alphabetical order,
Piero Basini,
C\'eline Blitz,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Ren\'e Matzen,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Tarje Nissen-Meyer,
Daniel Peter,
Max Rietmann,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Zhinan Xie,
Hejun Zhu.\\
%\toall{add full list of developers here and update the AUTHORS file accordingly}

@@

\item [Numerical simulations in general] ~\\
Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
Additional aspects dealing with adjoint simulations are described in
\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},

@@

The SCOTCH library \citep{PeRo96}
provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}. It is more recent than METIS, actively maintained and performs better in many cases.
A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
In case no SCOTCH libraries can be found on the system, the configuration will bundle this version for compilation.
The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
\end{lyxcode}

@@

\noindent
\texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp -qcache=auto -qhalt=w}\\
\noindent
\texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Wl,-relax -Q -Q+rank,swap\_all}.\\

@@

An efficient partitioning is important because it leverages the overall running time of the application.
It amounts to balance the number of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}. It is more recent than METIS, actively maintained and performs better in many cases.
A recent version of its source code is provided in directory \texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
\\

@@

Main authors: Dimitri Komatitsch and Jeroen Tromp

@@

 MPI v. 2.0, November 2010:
 Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
 Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
 and Emanuele Casarotti, INGV Roma, Italy:
  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
  much faster solver using Michel Deville's inlined matrix products.\\

>>>>

\section*{Authors}
The SPECFEM3D Cartesian package was first developed by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France 
from 1995 to 1997 and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.

Since then it has been developed and maintained by a development team: in alphabetical order,
Piero Basini,
C\'eline Blitz,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Dominik G\"oddeke,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Ren\'e Matzen,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Tarje Nissen-Meyer,
Daniel Peter,
Max Rietmann,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Jean-Pierre Vilotte,
Zhinan Xie,
Hejun Zhu.\\

@@

\item [Numerical simulations in general] ~\\
Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
Additional aspects of adjoint simulations are described in
\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},

@@

The SCOTCH library \citep{PeRo96}
provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
In case no SCOTCH libraries can be found on the system, the configuration will bundle the version provided with the source code for compilation.
The path to an existing SCOTCH installation can to be set explicitly with the option \texttt{-{}-with-scotch-dir}. Just as an example:
\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
\end{lyxcode}

@@

\noindent
\texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp -qcache=auto -qhalt=w}\\
\noindent
\texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Q -Q+rank,swap\_all -Wl,-relax}\\

@@

An efficient partitioning is important because it leverages the overall running time of the application.
It amounts to balance the number of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
\\

@@

Main historical authors: Dimitri Komatitsch and Jeroen Tromp

@@

 MPI v. 2.0, November 2010:
 Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
 Daniel Peter, Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
 and Emanuele Casarotti, INGV Roma, Italy:
  support for CUBIT meshes decomposed by SCOTCH;
  much faster solver using Michel Deville's inlined matrix products.\\
